---
title: "Homework 9"
author: "Kaleb Luna"
date: "2025-04-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
```

GitHub Link:

<P style="page-break-before: always">

\newpage

# Problem 1: Manufacturing flaws in circuit boards

## Part A

```{r echo = FALSE}
solderData <- read.csv("solder.csv")
# head(solderData)

ggplot(solderData, aes(x = Opening, y = skips)) + 
  geom_boxplot(fill = "maroon") + 
  labs(title = "Solder Skips vs. Opening Size", 
       x = "Opening Size", 
       y = "Number of Solder Skips",
       caption = "Solder skips tends to be more common as the opening size decreases.")

ggplot(solderData, aes(x = Solder, y = skips)) + 
  geom_boxplot(fill = "navy") + 
  labs(title = "Solder Skips vs. Solder Thickness", 
       x = "Solder Thickness",
       y = "Number of Solder Skips",
       caption = "Solder skips tend to be more common in thin solder types.")

```

## Part B

```{r echo = FALSE}
model <- lm(skips ~ Opening * Solder, data = solderData)

tidy_model <- tidy(model, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))
print(tidy_model)


```

The estimated coefficients and 95% confidence intervals for each respective variable and interaction:

Openings:

-   Large Opening (Intercept): 0.393, (-0.628, 1.415)

-   Medium Opening (OpeningM): 2.407, (0.962, 3.851)

-   Small Opening (OpeningS): 5.127, (3.682, 6.571)

Solders:

-   Thin Solder (Thin vs. Thick w/ Large Opening): 2.280, (0.836, 3.724)

-   Thin Solder with Medium Opening: -0.740, (-2.782, 1.302)

-   Thin Solder with Small Opening: 9.653, (7.611, 11.696)

## Part C

-   **Intercept:** With a large opening and a thick solder, we can predict that the number of solder skips is approximately 0.39. This estimate serves as the baseline factor for the following variables.

-   **OpeningM:** With a medium opening and a thick solder, we can predict that the number of solder skips increases by approximately 2.41.

-   **OpeningS:** With a small opening and a thick solder, we can predict that the number of solder skips increases by approximately 5.13.

-   **SolderThin:** With large opening and a thin solder, we can predict that the number of solder skips increases by approximately 2.28.

-   **OpeningM: SolderThin:** With a medium opening and a thin solder, there is approximately 0.74 fewer solder skips.

-   **OpeningS:SolderThin:** With a small opening and a thin solder, the number of solder skips increases by approximately 9.65.

## Part D

If I had to recommend a combination of Opening size and Solder thickness to AT&T based on the regression model, I would recommend the combination of a large opening and thick solder. This combination has the lowest predicted number of solder skips (0.39) compared to the rest of the combinations of openings and solders, and based on the graphs on Part A, larger opening sizes and thicker solders tend to have fewer solder skips.

<P style="page-break-before: always">

\newpage

# Problem 2: Grocery store prices

```{r echo = FALSE}
groceriesData <- read.csv("groceries.csv")
```

## Part A

```{r echo = FALSE}
storeAvg <- groceriesData |> 
  group_by(Store) |> 
  summarize(avg_price = mean(Price, na.rm = TRUE))

ggplot(storeAvg, aes(x = reorder(Store, avg_price), y = avg_price)) + 
  geom_col(fill = "lightblue") + 
  coord_flip() + 
  labs(title = "Average Product Price by Store",
       x = "Store",
       y = "Average Price in Dollars",
       caption = "The plot shows the average product price by store.")
```

Whole Foods has the highest average grocery prices in the data set, and Fiesta has the lowest average grocery prices. The graph above will later highlight how different outside factors (income, location, etc.) are related to average grocery prices.

## Part B

```{r echo = FALSE}
productCount <- groceriesData |> 
  group_by(Product) |> 
  summarize(num_stores = n_distinct(Store))

ggplot(productCount, aes(x = num_stores, y = reorder(Product, num_stores))) + 
  geom_col(fill = "maroon") + 
  labs(title = "Number of Stores Carrying Each Product",
       x = "Number of Stores", 
       y = "Product",
       caption = "The barplot shows the number of stores carrying each product.")
```

The bar plot above highlights how products vary across each store, specifically niche products such as health foods. In conclusion, this conveys how it may be misleading to conclude data based on average prices unless the availability of products is taken into account.

## Part C

(*Results of Regression Model Table Hidden due to Formatting Issues)*

```{r echo = FALSE, results = FALSE}
groceriesData$Type <- as.factor(groceriesData$Type)
groceriesData$Type <- relevel(groceriesData$Type, ref = "Grocery")

modelC <- lm(Price ~ Product + Type, data = groceriesData)
confint(modelC)
summary(modelC)

```

Compared with ordinary grocery stores (like Albertsons, HEB, or Krogers), convenience stores charge somewhere between 0.41 and 0.91 dollars more for the same product.

## Part D

```{r echo = FALSE, results = FALSE}
modelD <- lm(Price ~ Product + Store, data = groceriesData)
summary(modelD)

```

Walmart and Kroger Fresh Fare are the two stores that seem to charge the lowest prices when comparing the same product, approximately \$0.99 and \$0.90 than the baseline, respectively. In contrast, Wheatsville Food Co-Op and Whole Foods are the two stores that seem to charge the highest prices with respect to the same product.

## Part E

When analyzing and comparing the estimates of Central Market and H-E-B:

-   Central Market -0.57339

-   H-E-B: -0.64596

We can conclude that both stores are cheaper than the baseline store. However, H-E-B is slightly cheaper than Central Market by approximately 0.0726. Given this, it can be said that Central Market does charge relatively more than H-E-B for the same products, however, the difference is relatively small when comparing differences of other stores.

## Part F

```{r echo = FALSE, results = FALSE}
groceriesData <- groceriesData |> 
  mutate(Income10K = Income/10000)

modelF = lm(Price ~ Product + Income10K, data = groceriesData)
summary(modelF)

groceriesSDE <- groceriesData |> 
  mutate(
    zPrice = scale(Price),
    zIncome = scale(Income10K)
  )
modelFSDE <- lm(zPrice ~ Product + zIncome, data = groceriesSDE)
summary(modelFSDE)
```

When analyzing the Income10K coefficient:

-   -0.014090: negative

It can be concluded that consumers in poorer ZIP codes seem to pay more for the same product on average.

A one-standard deviation increase in the income of a ZIP code seems to be associated with a -0.0316 standard-deviation change in the price that consumers in that ZIP code expect to pay for the same product.

<P style="page-break-before: always">

\newpage

# Problem 3: redlining

## A. ZIP codes with a higher percentage of minority residents tend to have more FAIR policies per 100 housing units.

This statement is true. There is an apparent upward trend in Figure A1, and in regression model A, the positive coefficient of 0.014 and R\^2 is 0.52, indicating a positive trend between minority populations and FAIR policies.

## B. The evidence suggests an interaction effect between minority percentage and the age of the housing stock in the way that these two variables are related to the number of FAIR policies in a ZIP code.

This statement is undecidable. Model B only shows an interaction between minority percentage and age, with the coefficient for age being 0.398 with p = 0.125, indicating that it is not significant. There is no interaction that involves age, minority percentage in the FAIR model; therefore, the statement cannot be confirmed. To obtain the sufficient evidence, a model such as lm(polices \~ minority \* age, data = redlining) would be needed.

## C. The relationship between minority percentage and number of FAIR policies per 100 housing units is stronger in high-fire-risk ZIP codes than in low-fire-risk ZIP codes.

This statement is false. The coefficient of minority:fire_riskLow is -0.001, with a p-value of 0.839, meaning that it is not significant. The strength of the relationship is not significantly difference across both fire risk groups.

## D. Even without controlling for any other variables, income “explains away” all the association between minority percentage and FAIR policy uptake.

This statement is false. Even with income taken account, the minority percentage is still positively associated with FAIR policy uptake. In Model D1, the coefficient for minority percentage is 0.014, and in Model D2, the coefficient decreases to 0.01, however, it is still higher than the p of 0.002. This indicates that income does not "explain away" the relationship.

## E. Minority percentage and number of FAIR policies are still associated at the ZIP code level, even after controlling for income, fire risk, and housing age. 

This statement is true. With a minority coefficient of 0.008 and a p-value of 0.006, as well as the 95% CI (0.003, 0.014), it can be concluded that although the effect is smaller, the minority percentage is still significantly associated with the number of FAIR policies after controlling for income, fire risk, and housing age and remains statistically significant.
